|

### 2021-10-05 _main.Rmd
End of day:

Systematic reviews are a type of research that aim to use all existing evidence to provide the best answer to an important research question. They do this by finding and combining the results from many related primary research studies. Reviews involve multiple steps including: searching of existing studies; assessment of the studies against predefined inclusion criteria; collection of data from each study; assessment of each study's methods.

This chapter presents a systematic review of primary studies that have examined the relationship between the levels of blood lipids (such as cholesterol and triglycerides), and treatments that change these levels, and dementia.

There were XXX primary studies <!-- TODO check this -->that contained information on this relationship. I found that statins reduce the risk of Alzheimer's disease, but had no effect of vascular dementia. Lipids were not associated with any outcome. The methods used in some of the primary studies meant that I was less confident in the accuracy of their results.

@@

I sought to include studies that examine the relationship between blood lipid levels (or any specific lipid fraction, including total cholesterol, HDL, LDL, and triglycerides) and risk of incident dementia and its subtypes. Eligible study designs included randomized controlled trials and non-randomized observational studies of lipid modifying treatments, longitudinal studies examining the effect of increased/decreased blood lipid levels, and genetic instrumental variable (Mendelian randomization) studies examining the effect of genetically increased/decreased blood lipid levels.

Participants were screened for dementia at baseline and prevalent cases excluded. Alternatively, where no baseline screening was employed, participants were assumed to be dementia free if less than <50 years of age at baseline. Studies of any duration were included to allow for exploration of the effect of length of follow-up on the effect estimate using meta-regression. No limits were placed on the sample size of included studies.

Eligible studies defined dementia according to recognised criteria, for example the International Classification of Diseases (ICD),[@organizationwho1993] National Institute of Neurological Disorders and Stroke Association-Internationale pour la Recherche en l'Enseignement en Neurosciences (NINDS-AIREN),[@roman1993] or Diagnostic and Statistical Manual of Mental Disorders (DSM) criteria.[@edition2013] Studies utilising electronic health records were the exception to this, as it was assumed that these criteria were used when entering the outcome into the EHR.

@@

Inter- and intra-rater reliability during the screening stages were assessed for a 10% sub-sample of records. Intra-rater reliability involved a single reviewer applying the inclusion criteria to the same set of records while blinded to their previous decisions (i.e. assessment of conistency), while inter-rater reliability involved two reviewers independently screening the same set of records (i.e. assessment of accuracy).

Rater reliability was assessed using Gwet's agreement coefficient (AC1).[@gwet2008] This measure was chosen over other methods such as percent agreement (number of agreements divided by total number of assessments), as it accounts for chance agreement between reviewers but does not suffer from bias due to severely imbalanced marginal totals in the same way that Cohen's $kappa$ value does. [@cohen1960: @gwet2008; @wongpakaran2013] Given the small number of included studies in this review as a proportion of the total number screened, this is an important characteristic.

<!-- TODO Need to be sure of how to calculate. -->

@@

At present, no formalised risk-of-bias assessment tool for Mendelian randomization studies is available. Assessment of the risk of bias in Mendelian randomisation studies was informed by the approach used in a previous systematic review of Mendelian randomisation,[@mamluk2020] as identified by a review of risk-of-bias assessments in systematic reviews of Mendelian randomisation studies (advance results from this review were obtained from contact with the authors.). <!-- TODO CITATION NEEDED - add once published as a preprint --> A copy of this tool is available in Appendix \@ref(appendix-mr-rob), but in summary, results were assessed for bias arising from weak instruments, genetic and other confounding, pleiotropy, and selection of participants.

#### Risk of bias due to missing evidence

A recent shift towards the assessment of missing evidence due to selective non-reporting - as distinct from the selective reporting of a single result from multiple planned - is demonstrated via the forthcoming RoB-ME (Risk of Bias due to Missing Evidence in a synthesis) tool.[@zotero-15123] <!-- TODO Need more here describing missing evidence --> The tool is in development stages, and as part of this review, I piloted the tool, and provided feedback to the developers. <!-- TODO Cite Ben's work on effect of missing study versus choice of analysis method -->

@@

Both a fixed-effect and random-effects meta-analysis model was employed to combine the different included studies. <!-- TODO Need more information on the difference between the two approaches, and their formula here, and description of different methods and quarrels regarding distinction fixed effect vs random effects meta-analyses. --> The fixed-effect method was implemented as:

@@

\begin{equation}
  \theta_i = \mu + u_i
  (\#eq:meta-analysis-random)
\end{equation}

\begin{equation}
  weighted\ average = \frac{\sum Y_i (1/SE_i^2)}{\sum(1/SE_i^2)}
  (\#eq:meta-analysis-fixed)
\end{equation}

@@

#### Sensitivity analyses

I conducted a leave-one-out analysis in order to explore the impact of any results on the summary effect estimate. <!-- TODO CITATION NEEDED --> In addition, where there was evidence of heterogeneity between results included in a meta-analysis, I investigated this further using meta-regression against reported characteristics. _A priori_, I was interested in the effect that the age at baseline, sex and risk-of-bias judgement had on the results. <!-- TODO Need more here -->

Finally, I investigated the potential for small study effects, which may be caused by publication bias, both visually using funnel plots and statistically using Egger's regression test.  <!-- TODO CITATION NEEDED - cite Jonathon on interpreting small study effects, and whatever citation is needed for  -->

@@

#### Visualisation of results

Evidence maps are useful way to explore the distribution of research cohorts included in a systematic review.[@saran2018] As such, the location of each individual study contributing to the evidence base was quantified and visualised on a world map.

One of the limitations of current risk-of-bias assessments in systematic reviews is that they are often divorced from the results to which they refer, and are infrequently incorporated into the analysis.

In response to this criticism, I developed a new visualisation tool was designed to allow for "paired" forest plots, as recommended by the ROB2 publication, where the risk-of-bias assessment is presented alongside the results.[@sterne2019] This tool was developed as an adjunct to this thesis to aid in creating standardised risk-of-bias figures,[@mcguinness2020robvisPaper] and the "paired" forest plot functionality grew out of a collaboration with other researchers to design a modular method for creating custom forest plots.[@zotero-14999] <!-- Apparently collaboration is a big thing they want to see --> A summary of this tool is contained in Appendix \@ref(appendix-robvis), and all forest plots presented in this Chapter were created using this tool.

@@

__[Note: Julian, I am particularly interested in your feedback on this section, and the corresponding results section (Section \@ref(sys-rev-including-preprints)), as I am not convinced on the language I am using]__

Preprints are considered a valuable evidence source within this thesis (see Introduction, Section \@ref(diverse-sources-preprints)) but their inclusion in a systematic review.

As a sensitivity analysis, I explored the additional evidential value of including preprints in each meta-analysis performed, assessed using the fixed effect weight from a standard meta-analysis.

Additionally, I followed preprints up over time to investigate whether all identified preprints included in the review were subsequently published (in which case preprints provide a snapshot into the future, and a systematic review update would capture these reports) or alternatively, if some preprints were not published, then preprints provide a distinct evidence source.

@@

The database search identified `r num_to_text(23447)` records, of which

<!-- TODO Cross check this with the actual numbers included -->

Of the random sample of 500 records screened to ensure the accuracy of the study design filters, no eligible records were identified. Many of those excluded by the filters were basic science studies, commentaries or educational articles, as expected. <!-- TODO include some example citations here -->

@@

Following de-duplication, the titles and abstracts of `r num_to_text(16109)` records were assessed for eligibility. `r num_to_text(387)` were deemed potentially eligible, and the full text records for these were accessed and screened. <!-- TODO Check these figures & Finish PRISMA -->

A PRISMA (Preferred Reporting Items for Systematic Reviews and Meta-analyses) flow diagram,[@page2021] <!-- TODO CITATION NEEDED - cite PRIMSA flow paper here, if published --> presented in Figure \@ref(fig:prisma-flow-fig), illustrates the movement of articles through the review. To highlight the contribution of preprint archives to the review, the flow diagram delineates between those records captured through databases searches (presented on the right of the diagram) and those captured by the search tool described in the previous chapter (presented in grey on the left of the diagram).

Common reasons for exclusion at the full text-stage included studies that reported on the wrong exposures (n = XXX; most commonly a inelgible lipid fraction), used the wrong study design (n= XXX), or reported on a wrong outcome (n=24; e.g. change in cognitive scores).

<!-- TODO PRISMA must include reasons for exclusion - extract from Rayyan -->

@@

Those records which were excluded in the initial screening, but were included by the second reviewer (n=`r discrepancy_Inter`, Table \@ref(tab:agreeInter-table)) were investigated. This discrepancy between the two reviewers was explained in all cases by differing interpretations of the inclusion criteria, specifically around the definition of cognitive decline/MCI versus dementia, and the definition of eligible lipids fractions.

@@

### Characteristics of included studies

@@

Following full-text screening, 87 reports describing `r n_included` unique studies met the criteria for inclusion in the review.[@ancelin2012a; @ancelin2013; @andrews2019a; @arvanitakis2008; @batty2014; @benn2017b; @bettermann2012a; @beydoun2011; @bruce2017; @burgess2017; @chao2015a; @chen2014; @chiang2007; @chitnis2015; @chou2014a; @chuang2015; @cramer2008; @dodge2011; @forti2010; @gnjidic2016; @gottesman2017; @gustafson2012; @haag2009; @hayden2006; @heartprotectionstudycollaborativegroup2002; @hendrie2015; @hippisley-cox2010a; @jick2000; @kimm2011; @kivipelto2001; @kivipelto2005a; @kuo2015; @li2004; @li2005a; @li2010; @liao2013; @liu2019a; @mainousa.g.2005; @mielke2005a; @mielke2010a; @mielke2011; @mukherjee2013; @muller2007; @noale2013; @notkola1998; @ostergaard2017; @pan2018; @parikh2011; @peters2009; @raffaitin2009; @rantanen2017; @rea2005a; @redelmeier2019; @reitz2004a; @reitz2010; @ridker2008; @ronnemaa2011; @schilling2017a; @smeeth2009a; @so2017; @solomon2007; @solomon2009a; @solomon2010; @sparks2008; @strand2013; @svensson2019; @szwast2007; @tan2003a; @tynkkynen2016; @tynkkynen2018; @wang2012; @whitmer2005a; @yang2015; @yoshitake1995; @zamrini2004; @zandi2005; @zhu2018a; @zimetbaum1992] Table \@ref(tab:studyCharacteristics-table) presents a summary of the characteristics of each study.

The majority of studies were non-randomised studies with the sole two included randomised controlled trials both examining all-cause dementia. Similarly, a relatively small number of Mendelian randomisation studies were identified, and several performed a two-sample MR analysis using the same summary statistics from published GWAS, leading to complications in synthesis (see Section\@ref()).

Of the non-randomised studies examining treatments that modify lipid levels, the overwhelming majority examined statin use. For exposue studies, total cholesterol and LDL-cw were the most frequently reported lipid elements.

In terms of outcomes, the vast majority of studies examined either all-cause dementia or Alzheimer's disease, with only a small proportion examining vascular dementia (n=XXX, %). Some rarer outcome classifications such as vascular-component or mixed dementia were also investigated.

Of note, several studies provided evidence on a lipid fraction as part of a wider study on Mets. Raffatin/Ng

Many studies using electronic health records as their data source did not accurately report the diagnostic codes used to identify cases.

Midlife cholesterol levels were of particular interest, with several studies examining this age-range. <!-- TODO Talk a little bit more about age here -->

Three included reports were preprints (denoted in the Table \@ref(tab:studyCharacteristics-table) using an asterisk), one of which had subsequently been published and was captured by the primary literature search. All three included preprints were obtained from the bioRxiv preprint server and described Mendelian randomisation analyses.

Several Mendelian random studies made use of summary level data from the same published GWAS (the Global Lipids Genetics Consortium and the IGAP consortium <!-- TODO CITATION NEEDED -->). As this could incur double counting of participants, and so produce artifically precise estimates if meta-analysed, only results from one of these studies were included <!-- TODO How chosen?? -->. As discussed in Section \@ref(rev-discussion-MR), the choice was not material, as the results from all studies using these data sources were comparable.

<!--- TODO Figures and tables to include here:

-   Summary of risk of bias (not done)

--->

@@

\blandscape{}
<!----------------------------------------------------------------------------->
(ref:studyCharacteristics-caption) Characteristics of included studies, stratified by study design. Note that three studies reported on multiple study designs, and these have been duplicated across the relevant sub-sections.

@@

As illustrated in Figure \@ref(fig:cohortLocations), the majority of reports described studies conducted in high-income countries.

<!----------------------------------------------------------------------------->
```{r cohortLocationsSetup, include = FALSE}
```

@@

of interest, many of the included studies were peformed in Taiwan (`r n_taiwan`), all but one of which made use of the Taiwan National Health Insurance.

@@

As discussed, the results of the risk of bias assessments are presented alongside their corresponding result.

However, as a brief summary of the biases at play:

Both randomised controlled trials were considered to be at low risk of bias.

For NRSI
For NRSE

For MR studies

Additionally, Several synthesis were also assessed to be at risk of bias due to missing evidence. A key issue identified relating to bias due to missing data was the potential for the search to miss relevant Mendelian randomisation studies which had examined the association between Alzhiemer's disease and multiple risk factors. This is discussed further in Section \@ref(rev-discussion-MR).

<!-- TODO Discuss Taiwan health database as good evidence on non-Caucasian idenitiies -->

Haag is a good example of potentially missing results - number of cases of other outcoems reported, but no analysis performed.

### All-cause dementia

__Statins__

@@

@@

In contrast, a meta-analysis of prospective observational studies provided evidence of a protective effect of statins use on all-cause dementia risk (`r obsStatins$Dementia`, Figure \@ref(fig:obsStatinDementiaFig)).

@@

A single study examing the effect of HMGCR inhibition on any dementia, via a , and found weak evidence for an effect (`r mrStatin$Dementia`).

__Lipids__

* NRSE

* MR studies per 1-SD

In NRSI, statin use was associated with a reduce risk of all-cause dementia (`r obsStatins$Dementia`, Figure \@ref(fig:statinsNRSI)).

Several studies presented results per SD change in a lipid fraction.

In Mendelian randomisation studies, a similar

### Alzheimer's disease

There were no randomised trials of statin, or any other lipid regulating agent, use on Alzheimer's disease, though several observational studies reported on this outcome, providing evidence for a protective effect (`r obsStatins$AD`; Figure \@ref(fig:obsStatinADFig))

@@

Two Mendelian randomisation studies looked at specifically as a result of by HMGCR inhibition (mediated by a single SNP; rs17238484). The first used a one sample approach (SNP-exposure and SNP-outcome associations are estimated using the same dataset) in a large Copenhagen-based cohort, while the second made use of summary level data obtained from the Global Lipids Genetic Consortium (SNP-exposure) and the International Genomics of Alzheimer's Project (SNP-outcome).

>>>>

Systematic reviews are a type of research study that aim to collect and combine all existing evidence to provide the best possible answer to an important research question. Well-performed reviews involve multiple steps including: searching of existing studies; assessment of the studies against predefined inclusion criteria; collection of data from each study; and assessment of each study's methods.

This chapter presents a systematic review of primary studies that have examined the relationship between the levels of blood lipids (such as cholesterol and triglycerides), and treatments that change these levels such as statins, and dementia outcomes.

My review included `r n_included` primary studies that contained information on this relationship. I found that statins appear to reduce the risk of Alzheimer's disease, but had no effect of vascular dementia. Lipid levels were not associated with any outcome. The methods used in some of the primary studies meant that I was less confident in the accuracy of their results.

@@

I sought to include studies that examined blood lipid levels as a risk factor for demenita outcomes, defined either as binary hypercholesterolemia variable or by category/1-standard-deviation increase of a specific lipid fraction (total cholesterol, high- and low-density lipoprotein cholesterol and triglycerides). I also aimed to include studies examining the effect of treatments that modify lipids levels as a source of indirect evidence. Eligible study designs included randomized controlled trials and non-randomized observational studies of lipid modifying treatments, longitudinal studies examining the effect of increased/decreased blood lipid levels, and genetic instrumental variable (Mendelian randomization) studies examining the effect of genetically increased/decreased blood lipid levels.

Eligible studies screened participants for dementia at baseline and excluded any prevalent cases. Alternatively, where no baseline screening was employed, participants were assumed to be dementia free if less than 50 years of age at baseline. Studies of any duration were included to allow for exploration of the effect of length of follow-up on the effect estimate using meta-regression. No limits were placed on the sample size of included studies.

Eligible studies defined dementia outcomes according to recognised criteria, for example the International Classification of Diseases (ICD),[@organizationwho1993] National Institute of Neurological Disorders and Stroke Association-Internationale pour la Recherche en l'Enseignement en Neurosciences (NINDS-AIREN),[@roman1993] or Diagnostic and Statistical Manual of Mental Disorders (DSM) criteria.[@edition2013] Studies utilising electronic health records were the exception to this, as it was assumed that a valid criteria was employed when entering used when entering the outcome into the EHR.

Conference abstracts with no corresponding full-text publication were eligible, and where required, I contacted authors to obtain information on the study's status. No limitations were imposed on publication status, date, venue or language.

@@

Inter- and intra-rater reliability during the screening stages were assessed for a 10% sub-sample of records. Intra-rater reliability involved a single reviewer applying the inclusion criteria to the same set of records while blinded to their previous decisions (i.e. assessment of consistency), while inter-rater reliability involved two reviewers independently screening the same set of records (i.e. assessment of accuracy).

Rater reliability was assessed using Gwet's agreement coefficient (AC1).[@gwet2008] This measure was chosen over other methods such as percent agreement (number of agreements divided by total number of assessments), as it accounts for chance agreement between reviewers but does not suffer from bias due to severely imbalanced marginal totals in the same way that Cohen's $kappa$ value does. [@cohen1960: @gwet2008; @wongpakaran2013] Given the small number of included studies in this review as a proportion of the total number screened, this is a useful characteristic.

@@

At present, no formalised risk-of-bias assessment tool for Mendelian randomization studies is available. Assessment of the risk of bias in Mendelian randomisation studies was informed by the approach used in a previous systematic review of Mendelian randomisation,[@mamluk2020] as identified by a review of risk-of-bias assessments in systematic reviews of Mendelian randomisation studies (advance results from this review were obtained from contact with the review authors.). <!-- TODO CITATION NEEDED - add once published as a preprint --> A copy of this tool is available in Appendix \@ref(appendix-mr-rob), but in summary, results were assessed for bias arising from weak instruments, genetic and other confounding, pleiotropy and population stratification. Acceptable judgements for each fo the 5 domains in the tool included "low", "moderate" and "high" risk of bias.

#### Risk of bias due to missing evidence {#methods-rob-me}

In addition to assessing the risk of bias within each result contributing to a synthesis, I also assessed risk of bias due to missing evidence at the analysis level. This assessment examines evidence missing due to selective non-reporting - as distinct from the selective reporting of a single result from multiple planned - and was performed using the forthcoming RoB-ME (Risk of Bias due to Missing Evidence in a synthesis) tool.[@zotero-15123] <!-- TODO Need more here describing missing evidence --> The tool is in development stages, and as part of this review, I piloted the tool, and provided feedback to the developers. <!-- TODO Cite Ben's work on effect of missing study versus choice of analysis method -->

@@

For

A random-effects meta-analysis model was employed to combine the different included studies. <!-- TODO Need more information on the difference between the two approaches, and their formula here, and description of different methods and quarrels regarding distinction fixed effect vs random effects meta-analyses. --> The fixed-effect method was implemented as:

@@

Results were stratified into subgroups on the basis of the overall risk of bias assessment, and summary estimates for each subgroup, in addition to an overall effect estimate, are displayed. Additional descriptive statistics are presented, while prediction intervals are shown as a dotted line around the overall effect estimate. Finally, a test subgroup differences between studies at different levels of risk of bias was performed (see subsequent Section \@ref(sys-rev-visualising-results)).

Random-effects analyses assume that differences in study design and execution mean that the study effects are normally distributed around a common treatment effect ($\mu$) with a variance of $\tau^2$

\begin{equation}
  \theta_i \sim \mathcal{N}(\mu, \tau^2) (\#eq:eq9)
\end{equation}

<!-- Random-effects methods estimate $\tau^2$ to estimate the common effect ($\mu$). There are many estimators of $\tau^2$ with varying assumptions [@Veroniki2016; @Bakbergenuly2020], the details of which are beyond the scope of this article. The estimate of $\tau^2$ is incorporated into the weighting of each study -->

\begin{equation}
  w_i = \frac{1}{v_i + \hat{\tau^2}} (\#eq:eq10)
\end{equation}

<!-- A random-effects meta-analysis allows for study-specific effects, and assumes that the true effect estimated by each study varies and is assumed to be randomly and normally distributed on the transformed scale (e.g. log-ORs) between studies, with a variance of 𝜏2. The between-study variance is estimated (usually using the “method of moments” or DerSimonian and Laird estimate), and used to modify the weights of studies when calculating the summary effect estimate and variance. The weight of each study is 1𝑣𝑖+𝜏2 where 𝑣𝑖 is the variance of the 𝑖𝑡ℎ study and 𝜏2 is the estimate of the between-study variance. So long as the between-study variance is larger than 0, smaller studies will be given more weight in random-effects meta-analysis than in a fixed-effect meta-analysis. If the between-study variance is 0, then the random-effects model is equivalent to the fixed-effect model. -->
<!-- In both fixed-effect and random-effects inverse-variance meta-analysis, the summary effect estimate and variance are estimated in the same way. The effect estimates (𝜃𝑖) from each study can be (transformed) ORs, relative risks (RRs), standardised mean differences (SMDs) or other types of estimates: -->
<!-- 𝜃̂𝑆𝑢𝑚𝑚𝑎𝑟𝑦=⁡Σ𝑤𝑖𝜃𝑖Σ𝑤𝑖 -->
<!-- (2) -->
<!-- 𝑉𝑎𝑟𝑖𝑎𝑛𝑐𝑒⁡𝑜𝑓⁡𝜃̂𝑆𝑢𝑚𝑚𝑎𝑟𝑦=⁡1Σ𝑤𝑖 -->
<!-- (3) -->
<!-- where 𝜃𝑖 is effect estimate and 𝑤𝑖 is the weight from the 𝑖𝑡ℎ study on an approximately normal scale. -->

<!-- \begin{equation} -->
<!--   \theta_i = \mu + u_i -->
<!--   (\#eq:meta-analysis-random) -->
<!-- \end{equation} -->

<!-- \begin{equation} -->
<!--   $weighted\ average = \frac{\sum Y_i (1/SE_i^2)}{\sum(1/SE_i^2)}$ -->
<!--   (\#eq:meta-analysis-fixed) -->
<!-- \end{equation} -->

@@

#### Additional analyses

Where there was evidence of heterogeneity between results included in a meta-analysis, I investigated this further using meta-regression against reported characteristics. _A priori_, I was interested in the effect that the age at baseline, sex and risk-of-bias judgement had on the results. Syntheses with greater than 10 studies were assessed for heterogeneity across these covariates. <!-- TODO Need more here -->

Finally, I investigated the potential for small study effects, which may be caused by publication bias, both visually using funnel plots and formally using Egger's regression test. <!-- TODO CITATION NEEDED - cite Jonathon on interpreting small study effects, and whatever citation is needed for  -->

@@

#### Visualisation of results {#sys-rev-visualising-results}

Evidence maps are useful way to explore the distribution of research cohorts included in a systematic review.[@saran2018] As part of the initial descriptive synthesis, the location of each individual study contributing to the evidence base was quantified and visualised on a world map.

One of the limitations of current risk-of-bias assessments in systematic reviews is that they are often divorced from the results to which they refer, and are infrequently incorporated into the analysis. <!-- TODO CITATION NEEDED --> In response to this criticism, I developed a new visualisation tool was designed to allow for the production of "paired" forest plots, as recommended by the ROB2 publication, where a risk-of-bias assessment is presented alongside it's corresponding numerical results.[@sterne2019] This tool was developed as an adjunct to this thesis to aid in creating standardised risk-of-bias figures,[@mcguinness2020robvisPaper] and the "paired" forest plot functionality grew out of a collaboration with other researchers to design a modular method for creating custom forest plots.[@zotero-14999] <!-- Apparently collaboration is a big thing they want to see --> A summary of this tool is contained in Appendix \@ref(appendix-robvis), and all forest plots presented in this Chapter were created using this tool.

@@

_[Note: Julian, I am particularly interested in your feedback on this section, and the corresponding results section (Section \@ref(sys-rev-including-preprints)), as I am not convinced on the language I am using]_

Preprints are considered a valuable evidence source within this thesis (see Introduction, Section \@ref(diverse-sources-preprints)). As an adjunct analysis to this review, I explored the additional evidential value of including preprints in each meta-analysis performed, assessed using the fixed effect weight from a standard meta-analysis.

Additionally, I followed identified preprints up after a two-year lag to investigate whether they had been subsequently published (in which case preprints provide a snapshot into the future, and a systematic review update would capture these reports) or not (in which case preprints provide a distinct evidence source to conventional bibliographic databases).

@@

The database search identified `r num_to_text(23447)` records, of which `r comma(7338)` were duplicated records. Of the random sample of 500 records screened to ensure the accuracy of the study design filters, no eligible records were identified. Many of those excluded by the filters were basic science studies, commentaries or educational articles, as expected. <!-- TODO include some example citations here -->

@@

Following de-duplication, the titles and abstracts of `r num_to_text(16109)` records were assessed for eligibility. `r num_to_text(387)` were deemed potentially eligible, and the full text records for these were accessed and screened.

The PRISMA flow diagram <!-- TODO CITATION NEEDED - cite PRIMSA flow paper here, if published --> presented in Figure \@ref(fig:prisma-flow-fig), illustrates the movement of articles through the review. To highlight the contribution of preprint archives to the review, the flow diagram delineates between those records captured through databases searches (presented on the left of the diagram) and those captured by the search tool described in the previous chapter (presented in grey on the right of the diagram).

Common reasons for exclusion at the full text-stage included studies that reported on the wrong exposures (n = `r n_wrong_exposure` ; most commonly a ineligible lipid fraction), used the wrong study design (n= `r n_study_design`), or reported on a wrong outcome (n=24; e.g. change in cognitive scores).

@@

Those records which were excluded in the initial screening, but were included by the second reviewer (n=`r discrepancy_Inter`, Table \@ref(tab:agreeInter-table)) were investigated. This discrepancy between the two reviewers was explained in all cases by differing interpretations of the inclusion criteria, most commonly around the definition of cognitive decline versus dementia and the definition of eligible lipids fractions.

@@

### Characteristics of included studies {#sys-rev-characteristics}

@@

Following full-text screening, `r n_included` unique studies (described across `r n_reports_included` reports) met the criteria for inclusion in the review.`r get_all_citations()` Table \@ref(tab:studyCharacteristics-table) presents a summary of the characteristics of each study.

The majority of included studies described non-randomised analyses, with the sole two included randomised controlled trials (the Heart Protection Study/British Heart Foundation trial[@heartprotectionstudycollaborativegroup2002] and the JUPITER trial[@ridker2008]) both reporting on the effect of statin use on all-cause dementia in older adults. A similarly small number of Mendelian randomisation studies were identified, several of which employed a two-sample approach using the identical summary statistics from published genome wide association studies (GWAS) leading to complications in the synthesis (see Section \@ref(sys-rev-res-AD)).

Of the `r n_nrsi_included` non-randomised studies examining treatments that modify lipid levels, all examined statin use (`r n_statins`), while a small number also reported on other non-statins agents such as fibrates (`r n_fibrates`). In the `r n_nrse_included` non-randomised studies of exposure, hypercholesterolemia (`r n_hyperchol`) and total cholesterol levels (`r n_TC`) were the most frequently reported risk factors.

In terms of outcomes, the vast majority of studies examined either all-cause dementia (`r n_Dementia`) or Alzheimer's disease (`r n_AD`), with only a small proportion examining vascular dementia `r n_VaD`. Some rarer outcome classifications such as vascular-component or mixed dementia were also investigated, but were myc rarer.

<!-- Many studies using electronic health records as their data source did not accurately report the diagnostic codes used to identify cases. -->

<!-- Midlife cholesterol levels were of particular interest, with several studies examining this age-range.  -->

<!--- TODO Talk a little bit more about age here --->

Three included reports were preprints (denoted in the Table \@ref(tab:studyCharacteristics-table) using an asterisk),[@ridker2008; @so2017;@zhu2017] one of which had subsequently been published and was captured by the primary literature search.[@zhu2018] All three included preprints were obtained from the bioRxiv preprint server and all described a Mendelian randomisation analysis.

<!-- As discussed in Section \@ref(rev-discussion-MR), the choice was not material, as the results from all studies using these data sources were comparable. -->

<!-- TODO Figures and tables to include here: -   Summary of risk of bias (not done) -->

@@

\blandscape{}
<!----------------------------------------------------------------------------->
(ref:studyCharacteristics-caption) Characteristics of included studies, stratified by study design. Note that three studies reported on multiple analytical designs within a single study, and these have been duplicated across the relevant sub-sections.

@@

```{r cohortLocationsSetup, include = FALSE}
```

As illustrated in Figure \@ref(fig:cohortLocations), the majority of reports described studies conducted in high-income countries, with the most high represented region being North America. Of interest, several of the included studies were conducted in Taiwan (`r n_taiwan`), all but one of which made use of the Taiwan National Health Insurance database.

@@

Finally, there were several eligible studies reported as conference abstracts that did not present numerical results. These reports were included in the analysis to enable assessment of risk of bias due to missing evidence (see Section \@ref(methods-rob-me)).

@@

```{r riskOfBias, include = FALSE, message=FALSE, warning=FALSE}
```

As discussed above, the results of the risk-of-bias assessments are presented alongside their corresponding result. A more detailed discussion of the sources and directions of bias is presented in Chapter \@ref(tri-heading), and so this section presents a brief summary of the biases at play.

Following best practice, studies judged to be at critical risk of bias should be excluded from any quantitative analyses. <!-- TODO CITATION NEEDED --> `r num_to_words(n_critical,T)` observational studies were excluded on this basis, predominantly due to a lack of adjustment for any confounders (i.e. the study reported unadjusted estimates).`r critical_citations`

<!-- Both randomised controlled trials were considered to be at low risk of bias.  -->

<!-- For NRSI, common biases include immortal time (where participants in the exposure group are followed up) -->
<!-- For NRSE  -->

<!-- For MR studies, concerns over potential genetic pleiotropy -->

<!-- Additionally, several synthesis were also assessed to be at risk of bias due to missing evidence.  -->

<!-- A key issue identified relating to bias due to missing data was the potential for the search to miss relevant Mendelian randomisation studies which had examined the association between Alzhiemer's disease and multiple risk factors. This is discussed further in Section \@ref(rev-discussion-MR). -->

<!-- Several eligible studies included in the analysis did not have available results, either due to selective reporting of significant results in conference abstracts,[@su2017; @yamada2009b] or due to the unavailability of supplementary materials containing the relevant results.[@haag2009] -->

### All-cause dementia {#sys-rev-res-Dementia}

#### Statins

@@

```{r Hypercholesterolemia, include = FALSE, message=FALSE, warning=FALSE}
```

```{r mrLipidsAD, include = FALSE, message=FALSE, warning=FALSE}
```

```{r LipidsSD, include = FALSE, message=FALSE, warning=FALSE}
```

```{r lipidFrationsSetup, include = FALSE, message=FALSE, warning=FALSE}
```

@@

In contrast, a meta-analysis of `r obsStatins$Dementia$n` prospective observational studies provided some evidence of a protective effect of statins use on all-cause dementia risk (`r obsStatins$Dementia$estimate`, Figure \@ref(fig:obsStatinDementiaFig)).

@@

Finally, a single Mendelian randomisation analysis was identified examining the effect of  lowered LDL-c levels on the risk of all-cause dementia via genetic inhibition of the 3-hydroxy-3-methylglutaryl-coenzyme A reductase (HMGCR), emulating statin treatment (see Section \@ref(intro-statins) for more details of the statin mechanism of action). This analysis provided weak evidence for an effect (`r mrStatin$Dementia$estimate`).

&nbsp; <!---------------------------------------------------------------------->

#### Fibrates

```{r obsFibrates, include = FALSE, message=FALSE, warning=FALSE}
```

`r num_to_text(obsFibrates$Dementia$n,T)` studies examined the effect of fibrate use on all-cause dementia and found very weak evidence for an effect (`r obsFibrates$Dementia$estimate`, Figure \@ref(fig:obsFibrateDementiaFig)).

<!----------------------------------------------------------------------------->
(ref:obsFibrateDementia-cap) Random effects meta-analysis of non-randomised studies examining the effect of fibrate use on all-cause dementia

(ref:obsFibrateDementia-scap) Random effects meta-analysis of statins on all-cause dementia

```{r obsFibrateDementiaFig, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:obsFibrateDementia-cap)', out.width='100%', fig.scap='(ref:obsFibrateDementia-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/fp_obs_Fibrate_Dementia.png"))
```
<!----------------------------------------------------------------------------->

#### Lipids

Across all outcomes, lipid levels were categorised in a number of ways. The most common categorisation was hypercholesterolemia at baseline, defined most frequently as a total cholesterol measurement of greater than 6.5 mmol/L.

`r obsHyperchol$Dementia$n` studies reported on the association of hypercholesterolemia with all-cause dementia and provided weak evidence for an effect (`r obsHyperchol$Dementia$estimate`, Figure \@ref(fig:obsHyperDementia)).

<!----------------------------------------------------------------------------->
(ref:obsHyperDementia-cap) Random effects meta-analysis of non-randomised studies examining the effect of hypercholesterolemia on all-cause dementia

(ref:obsHyperDementia-scap) sMeta-analysis of hypercholesterolemia on all-cause dementia

```{r obsHyperDementia, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:obsHyperDementia-cap)', out.width='100%', fig.scap='(ref:obsHyperDementia-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/fp_obs_hyperchol_Dementia.png"))
```
<!----------------------------------------------------------------------------->

Several studies analysed individual lipid fractions by estimating the risk of dementia per 1 standard deviation increase in that fraction (Figure \@ref(fig:lipidFractionsDementia)). Very weak evidence for an effect on all-cause dementia was found for total cholesterol (`r n_effect(obsLipids$Dementia_TC)`), LDL-c (`r n_effect(obsLipids$Dementia_LDL)`), HDL-c (`r n_effect(obsLipids$Dementia_HDL)`) and triglycerides (`r n_effect(obsLipids$Dementia_TG)`).

<!----------------------------------------------------------------------------->
(ref:lipidFractionsDementia-cap) Random effects meta-analysis of four lipid fractions (total cholesteol, HDL, LDL, and triglycerides) on all-cause dementia risk, standardised per 1-SD increase in the lipid fraction.

(ref:lipidFractionsDementia-scap) Random effects meta-analysis of four lipid fractions on all-cause dementia
```{r lipidFractionsDementia, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:lipidFractionsDementia-cap)', out.width='100%', fig.scap='(ref:lipidFractionsDementia-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/fp_lipids_composite_Dementia.pdf"))
```
<!----------------------------------------------------------------------------->

Finally, there were no identified Mendelian randomisation analysis examining the effect of lower lipid levels, as determined by any genetic instrument, on all-cause dementia risk.

### Alzheimer's disease {#sys-rev-res-AD}

#### Statins

There were no randomised trials of the use of statins or any other lipid regulating agents  on Alzheimer's disease, though several observational studies reported on this outcome and provided evidence for a protective effect (`r n_effect(obsStatins$AD)`; Figure \@ref(fig:obsStatinADFig)).

@@

Two Mendelian randomisation studies looked at specifically as a result of HMGCR inhibition, mediated by SNPs in the gene (rs172338484 and rs12916).`r mrStatin$AD$citations` The first used a one sample approach (SNP-exposure and SNP-outcome associations are estimated using the same dataset) in a large Copenhagen-based cohort, while the second made use of summary level data obtained from the Global Lipids Genetic Consortium (SNP-exposure) and the International Genomics of Alzheimer's Project (SNP-outcome). Meta-analysis of these estimates provided weak evidence of an effect (`r mrStatin$AD$estimate`, Figure \@ref(fig:mrStatinADFig)).
